\begin{thebibliography}{99}

	\bibitem{church-turing} A.~Church, ``An unsolvable problem of elementary
	number theory,'' Amer.\ J.\ Math {\bf 58}, 345 (1936); A.~M.~Turing, ``On
	computable numbers, with an application to the Entscheidungsproblem,''
	Proc.\ Lond.\ Math.\ Soc.\ (2) {\bf 42}, 230 (1936) and Proc.\ Lond.\ Math.\
	Soc.\ (2) {\bf 43}, 544 (1937).
	\bibitem{feynman} R.~P.~Feynman, ``Simulating physics with computers,''
	Int.\ J.\ Theor.\ Phys.\ {\bf 21}, 467 (1982).
	\bibitem{shor-factoring} P.~Shor, ``Algorithms for quantum computation:
	discrete logarithms and factoring,'' Proceedings, 35th Annual Symposium
	on Fundamentals of Computer Science, (1994).
	\bibitem{grover} L.~K.~Grover, ``A fast quantum mechanical algorithm for
	database search,'' Proceedings, 28th ACM Symposium on Theory of Computation,
	212 (1996).
	\bibitem{bennett-strengths} C.~B.~Bennett, E.~Bernstein, G.~Brassard, and
	U.~Vazirani, ``Strengths and weaknesses of quantum computing,''
	quant-ph/9701001 (1997).
	\bibitem{cirac-zoller} J.~I.~Cirac and P.~Zoller, ``Quantum computations
	with cold trapped ions,'' Phys.\ Rev.\ Lett.\ {\bf 74}, 4091 (1995).
	\bibitem{wineland} C.~Monroe, D.~M.~Meekhof, B.~E.~King, W.~M.~Itano, and
	D.~J.~Wineland, ``Demonstration of a fundamental quantum logic gate,'' Phys.\
	Rev.\ Lett.\ {\bf 75}, 4714 (1995).
	\bibitem{kimble} Q.~A.~Turchette, C.~J.~Hood, W.~Lange, H.~Mabuchi, and
	H.~J.~Kimble, ``Measurement of conditional phase shifts for quantum logic,''
	Phys.\ Rev.\ Lett.\ {\bf 75}, 4710 (1995).
	\bibitem{gershenfeld} N.~Gershenfeld and I.~Chuang, ``Bulk spin resonance
	quantum computation,'' Science {\bf 275}, 350 (1997).
	\bibitem{shor-9qubit} P.~Shor, ``Scheme for reducing decoherence in quantum
	memory,'' Phys.\ Rev.\ A {\bf 52}, 2493 (1995).
	\bibitem{steane-7qubit} A.~M.~Steane, ``Error correcting codes in quantum
	theory,'' Phys.\ Rev.\ Lett.\ {\bf 77}, 793 (1996).
	\bibitem{cohen-tannoudji} C.~Cohen-Tannoudji, {\it Quantum Mechanics},
	Wiley, New York (1977).
	\bibitem{macwilliams-sloane} F.~J.~MacWilliams and N.~J.~A.~Sloane, {\it
	The Theory of Error-Correcting Codes}, North-Holland Publishing Company,
	New York (1977).
	\bibitem{no-cloning} W.~K.~Wootters and W.~H.~Zurek, ``A single quantum
	cannot be cloned,'' Nature {\bf 299}, 802 (1982).
	\bibitem{shannon} C.~E.~Shannon, ``A mathematical theory of communication,''
	Bell Sys.\ Tech.\ J.\ {\bf 27}, 379, 623 (1948).
	\bibitem{knill-laflamme-theory} E.~Knill and R.~Laflamme, ``A theory of
	quantum error-correcting codes,'' Phys.\ Rev.\ A {\bf 55}, 900 (1997).
	\bibitem{bennett-tome} C.~Bennett, D.~DiVincenzo, J.~Smolin, and W.~Wootters,
	``Mixed state entanglement and quantum error correction,'' Phys.\ Rev.\ A
		{\bf 54}, 3824 (1996).
	\bibitem{vaidman} L.~Vaidman, L.~Goldenberg, and S.~Wiesner, ``Error prevention
	scheme with four particles,'' Phys.\ Rev.\ A {\bf 54}, 1745R (1996).
	\bibitem{grassl} M.~Grassl, Th.~Beth, and T.~Pellizzari, ``Codes for the
	quantum erasure channel,'' quant-ph/9610042 (1996).
	\bibitem{leung} D.~W.~Leung, M.~A.~Nielsen, I.~L.~Chuang, Y.~Yamamoto,
	``Approximate quantum error correction can lead to better codes,''
	quant-ph/9704002 (1997).
	\bibitem{gottesman-stab} D.~Gottesman, ``Class of quantum error-correcting
	codes saturating the quantum Hamming bound,'' Phys.\ Rev.\ A {\bf 54}, 1862
	(1996).
	\bibitem{calderbank-stab} A.~R.~Calderbank, E.~M.~Rains, P.~W.~Shor, and
	N.~J.~A.~Sloane, ``Quantum error correction and orthogonal geometry,'' Phys.\
	Rev.\ Lett.\ {\bf 78}, 405 (1997).
	\bibitem{rains-shadow} E.~Rains, ``Quantum shadow enumerators,''
	quant-ph/9611001 (1996).
	\bibitem{laflamme-5qubit} R.~Laflamme, C.~Miquel, J.~P.~Paz, and W.~Zurek,
	``Pefect quantum error correction code,'' Phys.\ Rev.\ Lett.\ {\bf 77}, 198
	(1996).
	\bibitem{gottesman-pasting} D.~Gottesman, ``Pasting quantum codes,''
	quant-ph/9607027 (1996).
	\bibitem{calderbank-GF4} A.~R.~Calderbank, E.~M.~Rains, P.~W.~Shor, and
	N.~J.~A.~Sloane, ``Quantum error correction via codes over GF(4),''
	quant-ph/9608006 (1996).
	\bibitem{steane-8qubit} A.~Steane, ``Simple quantum error correcting codes,''
	Phys.\ Rev.\ A {\bf 54}, 4741 (1996).
	\bibitem{steane-RM} A.~Steane, ``Quantum Reed-Muller codes, ''
	quant-ph/9608026 (1996).
	\bibitem{calderbank-CSS} A.~R.~Calderbank and P.~W.~Shor, ``Good quantum
	error-correcting codes exist,'' Phys.\ Rev.\ A {\bf 54}, 1098 (1996).
	\bibitem{steane-CSS} A.~Steane, ``Multiple particle interference and quantum
	error correction,'' Proc.\ Roy.\ Soc.\ Lond.\ A {\bf 452}, 2551 (1996).
	\bibitem{knill-qudit} E.~Knill, ``Non-binary error bases and quantum codes,''
	quant-ph/9608048 (1996); E.~Knill, ``Group representations, error bases and
	quantum codes,'' quant-ph/9608049 (1996).
	\bibitem{chau-d^2} H.~F.~Chau, ``Correcting quantum errors in higher spin
	systems,'' quant-ph/9610023 (1996)
	\bibitem{chau-5qudit} H.~F.~Chau, ``Five quantum register error correction
	code for higher spin systems,'' quant-ph/9702033 (1997).
	\bibitem{aharonov} D.~Aharonov and M.~Ben-Or, ``Fault-tolerant quantum
	computation with constant error,'' quant-ph/9611025 (1996).
	\bibitem{rains-orthogonal} E.~Rains, ``Nonbinary quantum codes,''
	quant-ph/9703048 (1997).
	\bibitem{cleve} R.~Cleve and D.~Gottesman, ``Efficient computations of
	encodings for quantum error correction,'' quant-ph/9607030 (1996).
	\bibitem{divincenzo-decoder} D.~P.~DiVincenzo, ``Quantum gates and circuits,''
	quant-ph/9705009 (1997).
	\bibitem{shor-fault-tol} P.~Shor, ``Fault-tolerant quantum computation,''
	quant-ph/9605011 (1996).
	\bibitem{divincenzo} D.~DiVincenzo and P.~Shor, ``Fault-tolerant error
	correction with efficient quantum codes,'' Phys.\ Rev.\ Lett.\ {\bf 77},
	3260 (1996).
	\bibitem{gottesman-fault-tol} D.~Gottesman, ``A theory of fault-tolerant
	quantum computation,'' quant-ph/9702029 (1997).
	\bibitem{bennett-teleport} C.~H.~Bennett, G.~Brassard, C.~Crepeau, R.~Josza,
	A.~Peres, and W.~K.~Wootters, ``Teleporting an unknown quantum state via dual
	classical and Einstein-Podalsky-Rosen channels,'' Phys.\ Rev.\ Lett.\ {\bf 70},
	1895 (1993).
	\bibitem{bennett-EPP} C.~H.~Bennett, G.~Brassard, S.~Popescu, B.~Schumacher,
	J.~A.~Smolin, and W.~K.~Wootters, ``Purification of noisy entanglement and
	faithful teleportation via noisy channels,'' Phys.\ Rev.\ Lett.\ {\bf 76},
	722 (1996).
	\bibitem{gottesman-qudit} E.~Knill, R.~Laflamme, and D.~Gottesman, in
	preparation.
	\bibitem{knill-normalizer} E.~Knill, personal communication.
	\bibitem{knill-concatenate2} E.~Knill, R.~Laflamme, and W.~Zurek, ``Accuracy
	threshold for quantum computation,'' quant-ph/9610011 (1996); E.~Knill,
	R.~Laflamme, and W.~Zurek, ``Resilient quantum computation: error models
	and thresholds,'' quant-ph/9702058 (1997).
	\bibitem{evslin} J.~Evslin, S.~Kakade, and J.~P.~Preskill, unpublished.
	\bibitem{steane-correction} A.~M.~Steane, ``Active stabilization, quantum
	computation and quantum state synthesis,'' Phys.\ Rev.\ Lett.\ {\bf 78},
	2252 (1997).
	\bibitem{knill-concatenate1} E. Knill and R. Laflamme, ``Concatenated
	quantum codes,'' quant-ph/ 9608012 (1996).
	\bibitem{zalka} C.~Zalka, ``Threshold estimate for fault tolerant quantum
	computing,'' quant-ph/9612028 (1996).
	\bibitem{lloyd} S.~Lloyd, ``The capacity of a noisy quantum channel,''
	Phys.\ Rev.\ A {\bf 55}, 1613 (1997).
	\bibitem{schumacher} B.~Schumacher and M.~A.~Nielsen, ``Quantum data
	processing and error correction,'' Phys.\ Rev.\ A {\bf 54}, 2629 (1996).
	\bibitem{barnum} H.~Barnum, M.~A.~Nielsen, and B.~Schumacher, ``Information
	transmission through a noisy quantum channel,'' quant-ph/9702049 (1997).
	\bibitem{ekert} A.~Ekert and C.~Macchiavello, ``Error correction in quantum
	communication,'' Phys.\ Rev.\ Lett.\ {\bf 77}, 2585 (1996).
	\bibitem{cerf-cleve} N.~J.~Cerf and R.~Cleve, ``Information-theoretic
	interpretation of quantum error-correcting codes,'' quant-ph/9702031 (1997).
	\bibitem{shor-laflamme-QMW} P.~Shor and R.~Laflamme, ``Quantum analog of
	the MacWilliams identities for classical coding theory,'' Phys.\ Rev.\ Lett.\
		{\bf 78}, 1600 (1997).
	\bibitem{rains-enumerators} E.~M.~Rains, ``Quantum weight enumerators,''
	quant-ph/9612015 (1996).
	\bibitem{rains-poly-invariants} E.~M.~Rains, ``Polynomial invariants of quantum
	codes,'' quant-ph/9704042 (1997).
	\bibitem{cleve-classical} R.~Cleve, ``Quantum stabilizer codes and classical
	linear vodes,'' quant-ph/9612048 (1996).
	\bibitem{bennett-erasure} C.~H.~Bennett, D.~P.~DiVincenzo, and J.~A.~Smolin,
	``Capacities of quantum erasure channels,'' quant-ph/9701015 (1997).
	\bibitem{fuchs-KL} C.~Fuchs and J.~Smolin, unpublished.
	\bibitem{shor-smolin} P.~Shor and J.~Smolin, ``Quantum error-correcting codes
	need not completely reveal the error syndrome,'' quant-ph/9604006 (1996).
	\bibitem{rains-dist2} E.~M.~Rains, ``Quantum codes of minimum distance two,''
	quant-ph/ 9704043 (1997).
	\bibitem{rains-nonstab} E.~M.~Rains, R.~H.~Hardin, P.~W.~Shor, and
	N.~J.~A.~Sloane, ``A nonadditive quantum code,'' quant-ph/9703002 (1997).
	\bibitem{lloyd-universal} S.~Lloyd, ``Almost any quantum logic gate is
	universal,'' Phys.\ Rev.\ Lett.\ {\bf 75}, 346 (1995).
	\bibitem{gates} A.~Barenco, C.~H.~Bennett, R.~Cleve, D.~P.~DiVincenzo,
	N.~Margolus, P.~Shor, T.~Sleator, J.~Smolin, and H.~Weinfurter,
	``Elementary gates for quantum computation,'' Phys.\ Rev.\ A {\bf 52},
	3457 (1995).

\end{thebibliography}